The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2025-04-22 Paper.qmd
Minor bib and paper edits.

Although it is widely accepted that crying is a signal, there is no consensus on if or when it should be conceptualized as a minimal signal, a costly signal, an indexical signal, or a deceptive (manipulative) signal. Contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021], which would be type of minimal signal. Demonstrating vigor and health early in life was likely vital for much of our evolutionary history due to the risk of neglect or infanticide experienced by children with poor survival prospects [@daly1984; @soltis2004]. As crying comes with physiological costs [@rao1997; @thureen1998], it could serve as a credible signal of vigor if these costs are too high for unhealthy infants to bear [@furlow1997; @lummaa1998; @soltis2004; @wells2003]. Children with severe medical conditions also produce acoustically distinct cries, suggesting that some elements of crying are indexical [@soltis2004]. Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @zahavi1977a; @eshel1991; @bergstrom1999 @kennedy2021].

@@

<!-- [One dense paragraph on how what we have learned relates to what has been reported empirically in the past]
- Note: we did not mention declines with age in the introduction
- Vingerhoets' lab and their colleagues maintain crying is understudied with older children

[Move first sentence down to 468 and broaden? Or at least broaden? MacFarlane's study of tantrums among Berkeley: declines from 5 to 12]

The cries and fussing of young infants increases until 6 weeks, peaking in some populations at over two hours a day, and then declines steadily until 12-16 weeks [@vermillet2022; @barr1991].
-->

@@

@@

@@

## Signal frequency vs. type

>>>>

Although it is widely accepted that crying is a signal, there is no consensus on if or when it should be conceptualized as a minimal signal, a costly signal, an indexical signal, or a deceptive (manipulative) signal. Contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021], which would be type of minimal signal. Demonstrating vigor and health early in life was likely vital for much of our evolutionary history due to the risk of neglect or infanticide experienced by children with poor survival prospects [@daly1984; @soltis2004]. As crying comes with physiological costs [@rao1997; @thureen1998], it could serve as a credible signal of vigor if these costs are too high for unhealthy infants to bear [@furlow1997; @lummaa1998; @soltis2004; @wells2003]. Children with severe medical conditions also produce acoustically distinct cries, suggesting that some elements of crying are indexical [@soltis2004]. Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @zahavi1977a; @lummaa1998; @eshel1991; @bergstrom1999 @kennedy2021].

@@

@@

["perceptions of need" instead of "perceived higher"?]

@@

[new additions above too repetive?]

@@
